Learning From Massive Noisy Labeled Data for Image Classification

### 2015-04-09 paper.tex
Polish by xgwang

%%%%%%%%% ABSTRACT
\begin{abstract}
Large-scale supervised datasets are crucial to train convolutional neural networks (CNNs) for various computer vision problems. However, obtaining a massive amount of well labeled data is usually very expensive and time consuming. In this paper, we introduce a general framework to train CNNs with only limited number of clean labels and millions of easily obtained noisy labels. We describe label noises with a probabilistic graphical model and further integrate it into an end-to-end deep learning system. To demonstrate the effectiveness of our approach, we collect a large-scale clothing classification dataset with both noisy and clean labels. Experiments on this dataset show that our model can better detect and correct the wrong labels, which benefits the training of underlying CNNs.
\end{abstract}

@@

Deep learning with a massive amount of supervised training data has recently shown very impressive improvement on multiple image recognition challenges including image classification~\cite{krizhevsky2012imagenet}, attribute learning~\cite{zhang2013panda}, scene classification~\cite{farabet2013learning}, \etc. While state-of-the-art results have been continuously reported~\cite{zeiler2013visualizing,simonyan2014very,szegedy2014going}, all these methods require reliable annotations from millions of images~\cite{deng2009imagenet} which are often expensive and time-consuming to obtain~\cite{deng2009imagenet}, preventing deep models from being quickly trained on new image recognition problems. Thus it is necessary to develop new efficient labeling and training frameworks for deep learning.

One possible solution is to automatically collect large amount of annotations from the Internet web images~\cite{gong2014multiview} (i.e. extracting tags from the surrounding texts or keywords from search engines) and directly use them as ground truth to train deep models. Unfortunately, these labels are extremely unreliable due to the various types of noise (i.e. labeling mistakes from annotators or computing errors from extraction algorithms). Many works have shown that these noisy labels could adversely impact the classification accuracy of the induced classifiers~\cite{zhu2004class,nettleton2010study,pechenizkiy2006class}. Various label noise-robust algorithms are developed but experiments show that performances of classifiers inferred by robust algorithms are still affected by label noise~\cite{bartlett2006convexity,teng2001comparison}. Other data cleansing algorithms are proposed~\cite{barandela2000decontamination,brodley2011identifying,miranda2009use}, but these approaches are difficult in distinguishing informative hard examples from harmful mislabeled ones.

@@

Although annotating all the data is costly, it is often easy to obtain a small amount of clean labels. Based on the observation of transferability of deep neural networks, people first initialize the parameters with a model pretrained on a larger yet related dataset~\cite{krizhevsky2012imagenet}, and then finetune on the smaller dataset of specific tasks~\cite{oquab2013learning,azizpour2014generic,donahue2013decaf}. Such methods may better avoid overfitting and utilize the relationships between the two datasets. However, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it only on the clean labels. Other approaches address the problem as semi-supervised learning where noisy labels are discarded~\cite{zhu2002learning}. These algorithms usually suffer from model complexity thus cannot be applied on large-scale datasets. Therefore, it is inevitable to develop a better way of using huge amount of noisy labeled data.

Our goal is to build an end-to-end deep learning system that is capable of training with both limited clean labels and massive noisy labels more effectively. Figure~\ref{fig:overview} shows the framework of our approach. We collect $1,000,000$ clothing images from some online shopping websites. Each image is automatically assigned with a noisy label according to the keywords in its surrounding text. After that, we manually refine $72,409$ image labels, which constitute a clean sub-dataset. All the data are then used to train CNNs, while the major challenge is to identify and correct the wrong labels during the training process.

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the true labels and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modifies a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumes noisy labels are conditionally independent with input images given clean labels. By examining our collected dataset, we find that this assumption is too strong to fit the real-world data well. For example, in Figure~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. It shows that images tend to be mislabeled may share similar visual patterns. This is because different sellers have their own bias on different categories, thus they may provide wrong keywords for many similar clothes. Based on these observations, we introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Figure~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or false conversion from the text to label (\eg, the samples with ``$\times$'' in Figure~\ref{fig:overview}).
\end{itemize}

The proposed probabilistic model builds the relations among the image, noisy label, ground truth label, and the noise type, where the latter two are treated as latent variables. We use the Expectation-Maximization (EM) algorithm to solve the problem and integrate it into the training process of CNNs. Experiments on the collected dataset show that our model can better detect and correct the wrong labels.

@@

\section{Related Work} % (fold)
\label{sec:related_work}
For most of the related works including effect of label noises, taxonomy of label noises, robust algorithms and noise cleaning algorithms for learning with noisy data, we refer to~\cite{frenay2013classification} for a comprehensive review.

{\bf Direct learning with noisy labels:} Many works have shown that label noise can adversely impact the classification accuracy of induced classifiers~\cite{zhu2004class}. To better handle label noise, some approaches rely on training classifiers with label noise-robust algorithms~\cite{beigman2009learning,manwani2013noise}. However,~\cite{bartlett2006convexity} proves that most of the loss functions are not completely robust to label noise. Experiments in~\cite{teng2001comparison} show that the classifiers inferred by label noise-robust algorithms are still affected by label noise. These methods seem to be adequate only when label noise can be safely managed by overfitting avoidance~\cite{frenay2013classification}. On the other hand, some label noise cleansing methods are proposed to remove or correct mislabeled instances~\cite{barandela2000decontamination,brodley2011identifying,miranda2009use}, but these approaches are difficult in distinguishing informative hard examples from harmful mislabeled ones. Thus they may remove too many instances and the overcleansing can reduce the performances of classifiers~\cite{matic1992computer}.

{\bf Semi-supervised learning:} Apart from directly learning with label noise, some semi-supervised learning algorithms are developed to utilize weakly labeled or even unlabeled data. Label Propagation method~\cite{zhu2002learning} explicitly uses ground truths of well labeled data to classify unlabeled samples. However, it suffers from computing pairwise distance, which has quadratic complexity with the number of samples thus cannot be applied on large-scale datasets. Chen \etal~\cite{chen2013fast} tackled this problem in the scenario of image tagging by a co-regularized duo classifier that can efficiently use both the image and incomplete tags. Weston \etal~\cite{weston2012deep} proposed to embed a pairwise loss in the middle layer of a deep neural network, which benefits the learning of discriminative features. But they need extra information about whether a pair of unlabeled images belong to the same class, which cannot be obtained in our problem.

@@

{\bf Noise modeling with deep learning:} Various methods have been proposed to handle label noise in different problem settings, but there are very few works about deep learning from noisy labels~\cite{mnih2012learning, larsen1998design, sukhbaatar2014learning}.~\cite{mnih2012learning} builds a simple noise model for aerial images but only considers binary classification.~\cite{larsen1998design} assumes label noise is independent from the true class label which is a simple and special case.~\cite{sukhbaatar2014learning} generalizes from them by considering multi-class classification and modeling class dependent noise, but they assume the noise is conditionally independent with the image content, ignoring the hardness of labeling images of different confusing levels. Our work can be viewed as a generalization of~\cite{sukhbaatar2014learning,natarajan2013learning} and our model is flexible enough to not only class dependent noise but also input dependent noise.
% section related_work (end)

\section{Label Noise Model} % (fold)
\label{sec:label_noise_model}
We target on learning a classifier from a set of images with noisy labels. To be specific, we have a noisy labeled dataset $\mathcal{D}_n=\left\{\left(\vec{x}^{(1)}, \tilde{y}^{(1)}\right), \dots, \left(\vec{x}^{(N)}, \tilde{y}^{(N)}\right)\right\}$ with the $n$-th image $\vec{x}^{(n)}$ and its corresponding noisy label $\tilde{y}^{(n)} \in \{1,\dots,L\}$, where $L$ is the number of classes. We describe how the noisy label is generated by using a probabilistic graphical model shown in Figure~\ref{fig:noise_pgm}.

@@

On the other hand, the label noise type $\vec{z}$ is an 1-of-3 binary random variable. It is associated with three different semantic meanings:
\begin{enumerate}
    \item The label is noise free, \ie, $\tilde{\vec{y}}$ should be equal to $\vec{y}$
    \item The label suffers from a pure random noise, \ie, $\tilde{\vec{y}}$ can take any possible value other than $\vec{y}$
    \item The label suffers from a confusing noise, \ie, $\tilde{\vec{y}}$ can take several values that are confusing with $\vec{y}$
\end{enumerate}

Following this assignment rule, we define the conditional probability of the noisy label as
\begin{equation} \label{eq:likelihood}
    p(\tilde{\vec{y}}|\vec{y},\vec{z}) = \begin{cases}
        \tilde{\vec{y}}^{T}\mat{I}\vec{y} & \text{if } \vec{z}_1 = 1 \\
        \frac{1}{L-1}\tilde{\vec{y}}^{T}(\mat{U} - \mat{I})\vec{y} & \text{if } \vec{z}_2 = 1 \\
        \tilde{\vec{y}}^{T}\mat{C}\vec{y} & \text{if } \vec{z}_3 = 1
    \end{cases}
\end{equation}
where $\mat{I}$ is the identity matrix, $\mat{U}$ is the unit matrix (all the elements are one), and $\mat{C}$ is a sparse stochastic matrix with $\mathrm{tr}(\mat{C})=0$ and $\mat{C}_{ij}$ denoting the confusion probability between class $i$ and $j$. Then we can derive from Figure~\ref{fig:noise_pgm} the joint distribution of $\tilde{\vec{y}}, \vec{y}$ and $\vec{z}$ conditioning on $\vec{x}$:
\begin{equation} \label{eq:joint_distribution}
  p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}) = p(\tilde{\vec{y}} | \vec{y}, \vec{z}) p(\vec{y} | \vec{x}) p(\vec{z} | \vec{x})
\end{equation}

While the class label probability distribution $p(\vec{y} | \vec{x})$ is comprehensible, the semantic meaning of $p(\vec{z} | \vec{x})$ needs extra clarification: it represents how confusing the image content is. Specific to our clothing classification problem, $p(\vec{z} | \vec{x})$ can be affected by different factors, including background clutter, image resolution, the style and material of the clothes, \etc. We show some examples in Figure~\ref{fig:compare_noise_level}.

@@

To illustrate the relations between noisy label and ground truth, we can derive their conditional probability from Eq~\ref{eq:joint_distribution}
\begin{equation}
    p(\tilde{\vec{y}} | \vec{y}, \vec{x}) = \sum_{\vec{z}} p(\tilde{\vec{y}}, \vec{z} | \vec{y}, \vec{x}) = \sum_{\vec{z}} p(\tilde{\vec{y}} | \vec{y}, \vec{z}) p(\vec{z} | \vec{x})
\end{equation}
which can be interpreted as a mixture model. Given an input image $\vec{x}$, the conditional probability $p(\vec{z} | \vec{x})$ can be seen as the prior of each mixture component. This makes a key difference between our work and~\cite{sukhbaatar2014learning}, where they assume $\tilde{\vec{y}}$ is conditionally independent with $\vec{x}$ if $\vec{y}$ is given. All the images share a same noise model in~\cite{sukhbaatar2014learning}, while in our approach each data sample has its own.

\subsection{Learning the Parameters} % (fold)
\label{sub:learning_the_parameters}
We exploit two CNNs to model $p(\vec{y} | \vec{x})$ and $p(\vec{z} | \vec{x})$ separately. Denote the parameter set of each deep model by $\theta_1$ and $\theta_2$. Our goal is to find the optimal $\theta = \theta_1 \cup \theta_2$ that maximize the incomplete log-likelihood $\log p(\tilde{\vec{y}} | \vec{x}; \theta)$. The Expectation-Maximization algorithm is used to iteratively solve this problem.

For any probability distribution $q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x})$, we can derive a lower bound of the incomplete log-likelihood
\begin{equation}
\begin{aligned}
    \log p(\tilde{\vec{y}} | \vec{x}; \theta) &= \log \sum_{\vec{y},\vec{z}} p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta) \\
    &\geq \sum_{\vec{y},\vec{z}} q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) \log \frac{p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta)}{q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x})}
\end{aligned}
\end{equation}
\textbf{E-Step} The difference between $\log p(\tilde{\vec{y}} | \vec{x}; \theta)$ and its lower bound is the Kullback-Leibler divergence $\mathrm{KL}\left(q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) || p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta)\right)$, which is equal to zero if and only if $q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) = p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta)$. Therefore, in each iteration $t$, we first compute the posterior of latent variables using current parameters $\theta^{(t)}$:
\begin{multline} \label{eq:posterior}
    p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) = \frac{p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta^{(t)})}{p(\tilde{\vec{y}} | \vec{x}; \theta^{(t)})} \\
    = \frac{p(\tilde{\vec{y}}|\vec{y},\vec{z};\theta^{(t)}) p(\vec{y}|\vec{x};\theta^{(t)}) p(\vec{z}|\vec{x};\theta^{(t)})}{\sum_{\vec{y}',\vec{z}'} p(\tilde{\vec{y}}|\vec{y}',\vec{z}';\theta^{(t)}) p(\vec{y}'|\vec{x};\theta^{(t)}) p(\vec{z}'|\vec{x};\theta^{(t)})}
\end{multline}
Then the expected complete log-likelihood can be written as
\begin{equation}
    Q(\theta; \theta^{(t)}) = \sum_{\vec{y},\vec{z}} p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \log p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta)
\end{equation}
\textbf{M-Step} We exploit two CNNs to model the probability $p(\vec{y}|\vec{x};\theta_1)$ and $p(\vec{z}|\vec{x};\theta_2)$, respectively. Thus the gradient of $Q$ \wrt $\theta$ can be decoupled into two parts:
\begin{multline} \label{eq:gradient}
    \frac{\partial Q}{\partial \theta} = \sum_{\vec{y},\vec{z}} p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta} \log p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta) \\
    = \sum_{\vec{y}} p(\vec{y} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta_1} \log p(\vec{y}|\vec{x};\theta_1) + \\
      \sum_{\vec{z}} p(\vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta_2} \log p(\vec{z}|\vec{x};\theta_2)
\end{multline}

The M-Step above is equivalent to minimize the cross entropy between the estimated ground truth distribution and the prediction of the classifier.
% subsection learning_the_parameters (end)

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
Large-scale supervised datasets are crucial to train convolutional neural networks (CNNs) for various computer vision problems. However, obtaining a massive amount of well labeled data is usually very expensive and time consuming. In this paper, we introduce a general framework to train CNNs with only a limited number of clean labels and millions of easily obtained noisy labels. We describe label noises with a probabilistic graphical model and further integrate it into an end-to-end deep learning system. To demonstrate the effectiveness of our approach, we collect a large-scale clothing classification dataset with both noisy and clean labels. Experiments on this dataset show that our model can better detect and correct the wrong labels, which benefits the training of underlying CNNs.
\end{abstract}

@@

Deep learning with a massive amount of supervised training data has recently shown very impressive improvement on multiple image recognition challenges including image classification~\cite{krizhevsky2012imagenet}, attribute learning~\cite{zhang2013panda}, and scene classification~\cite{farabet2013learning}. While state-of-the-art results have been continuously reported~\cite{zeiler2013visualizing,simonyan2014very,szegedy2014going}, all these methods require reliable annotations from millions of images~\cite{deng2009imagenet} which are often expensive and time-consuming to obtain~\cite{deng2009imagenet}, preventing deep models from being quickly trained on new image recognition problems. Thus it is necessary to develop new efficient labeling and training frameworks for deep learning.

One possible solution is to automatically collect large amount of annotations from the Internet web images~\cite{gong2014multiview} (i.e. extracting tags from the surrounding texts or keywords from search engines) and directly use them as ground truth to train deep models. Unfortunately, these labels are extremely unreliable due to various types of noise (i.e. labeling mistakes from annotators or computing errors from extraction algorithms). Many works have shown that these noisy labels could adversely impact the classification accuracy of the induced classifiers~\cite{zhu2004class,nettleton2010study,pechenizkiy2006class}. Various label noise-robust algorithms are developed but experiments show that performances of classifiers inferred by robust algorithms are still affected by label noise~\cite{bartlett2006convexity,teng2001comparison}. Other data cleansing algorithms are proposed~\cite{barandela2000decontamination,brodley2011identifying,miranda2009use}, but these approaches are difficult in distinguishing informative hard examples from harmful mislabeled ones.

@@

Although annotating all the data is costly, it is often easy to obtain a small amount of clean labels. Based on the observation of transferability of deep neural networks, people initialize parameters with a model pretrained on a larger yet related dataset~\cite{krizhevsky2012imagenet}, and then finetune on the smaller dataset of specific tasks~\cite{oquab2013learning,azizpour2014generic,donahue2013decaf}. Such methods may better avoid overfitting and utilize the relationships between the two datasets. However, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it only on clean labels. Other approaches address the problem as semi-supervised learning where noisy labels are discarded~\cite{zhu2002learning}. These algorithms usually suffer from model complexity thus cannot be applied on large-scale datasets. Therefore, it is inevitable to develop a better way of using huge amount of noisy labeled data.

Our goal is to build an end-to-end deep learning system that is capable of training with both limited clean labels and massive noisy labels more effectively. Figure~\ref{fig:overview} shows the framework of our approach. We collect $1,000,000$ clothing images from online shopping websites. Each image is automatically assigned with a noisy label according to the keywords in its surrounding text. We manually refine $72,409$ image labels, which constitute a clean sub-dataset. All the data are then used to train CNNs, while the major challenge is to identify and correct wrong labels during the training process.

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the true labels and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modified a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumed noisy labels were conditionally independent with input images given clean labels. By examining our collected dataset, we find that this assumption is too strong to fit real-world data well. For example, in Figure~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. It shows that images tend to be mislabeled may share similar visual patterns. This is because different sellers have their own bias on different categories, thus they may provide wrong keywords for many similar clothes. Based on these observations, we introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Figure~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or false conversion from the text to label (\eg, the samples with ``$\times$'' in Figure~\ref{fig:overview}).
\end{itemize}

The proposed probabilistic model builds the relations among images, noisy labels, ground truth labels, and noise types, where the latter two are treated as latent variables. We use the Expectation-Maximization (EM) algorithm to solve the problem and integrate it into the training process of CNNs. Experiments on the collected dataset show that our model can better detect and correct the wrong labels.

@@

\section{Related Work} % (fold)
\label{sec:related_work}
For most of the related works including the effect of label noises, taxonomy of label noises, robust algorithms and noise cleaning algorithms for learning with noisy data, we refer to~\cite{frenay2013classification} for a comprehensive review.

{\bf Direct learning with noisy labels:} Many works have shown that label noise can adversely impact the classification accuracy of induced classifiers~\cite{zhu2004class}. To better handle label noise, some approaches rely on training classifiers with label noise-robust algorithms~\cite{beigman2009learning,manwani2013noise}. However, Bartlett \etal~\cite{bartlett2006convexity} proved that most of the loss functions are not completely robust to label noise. Experiments in~\cite{teng2001comparison} showed that the classifiers inferred by label noise-robust algorithms are still affected by label noise. These methods seem to be adequate only when label noise can be safely managed by overfitting avoidance~\cite{frenay2013classification}. On the other hand, some label noise cleansing methods were proposed to remove or correct mislabeled instances~\cite{barandela2000decontamination,brodley2011identifying,miranda2009use}, but these approaches were difficult in distinguishing informative hard examples from harmful mislabeled ones. Thus they might remove too many instances and the overcleansing could reduce the performances of classifiers~\cite{matic1992computer}.

{\bf Semi-supervised learning:} Apart from direct learning with label noise, some semi-supervised learning algorithms were developed to utilize weakly labeled or even unlabeled data. The Label Propagation method~\cite{zhu2002learning} explicitly used ground truths of well labeled data to classify unlabeled samples. However, it suffered from computing pairwise distance, which has quadratic complexity with the number of samples thus cannot be applied on large-scale datasets. Weston \etal~\cite{weston2012deep} proposed to embed a pairwise loss in the middle layer of a deep neural network, which benefits the learning of discriminative features. But they needed extra information about whether a pair of unlabeled images belong to the same class, which cannot be obtained in our problem.

@@

{\bf Noise modeling with deep learning:} Various methods have been proposed to handle label noise in different problem settings, but there are very few works about deep learning from noisy labels~\cite{mnih2012learning, larsen1998design, sukhbaatar2014learning}. Mnih and Hinton~\cite{mnih2012learning} built a simple noise model for aerial images but only considered binary classification. Larsen \etal~\cite{larsen1998design} assumed label noises are independent from true class labels which is a simple and special case. Sukhbaatar \etal~\cite{sukhbaatar2014learning} generalized from them by considering multi-class classification and modeling class dependent noise, but they assumed the noise was conditionally independent with the image content, ignoring the hardness of labeling images of different confusing levels. Our work can be viewed as a generalization of~\cite{sukhbaatar2014learning,natarajan2013learning} and our model is flexible enough to not only class dependent noise but also input dependent noise.
% section related_work (end)

\section{Label Noise Model} % (fold)
\label{sec:label_noise_model}
We target on learning a classifier from a set of images with noisy labels. To be specific, we have a noisy labeled dataset $\mathcal{D}_{\eta}=\left\{\left(\vec{x}^{(1)}, \tilde{y}^{(1)}\right), \dots, \left(\vec{x}^{(N)}, \tilde{y}^{(N)}\right)\right\}$ with $n$-th image $\vec{x}^{(n)}$ and its corresponding noisy label $\tilde{y}^{(n)} \in \{1,\dots,L\}$, where $L$ is the number of classes. We describe how the noisy label is generated by using a probabilistic graphical model shown in Figure~\ref{fig:noise_pgm}.

@@

The label noise type $\vec{z}$ is an 1-of-3 binary random variable. It is associated with three different semantic meanings:
\begin{enumerate}
    \item The label is noise free, \ie, $\tilde{\vec{y}}$ should be equal to $\vec{y}$
    \item The label suffers from a pure random noise, \ie, $\tilde{\vec{y}}$ can take any possible value other than $\vec{y}$
    \item The label suffers from a confusing noise, \ie, $\tilde{\vec{y}}$ can take several values that are confusing with $\vec{y}$
\end{enumerate}

Following this assignment rule, we define the conditional probability of the noisy label as
\begin{equation} \label{eq:likelihood}
    p(\tilde{\vec{y}}|\vec{y},\vec{z}) = \begin{cases}
        \tilde{\vec{y}}^{T}\mat{I}\vec{y} & \text{if } \vec{z}_1 = 1 \\
        \frac{1}{L-1}\tilde{\vec{y}}^{T}(\mat{U} - \mat{I})\vec{y} & \text{if } \vec{z}_2 = 1 \\
        \tilde{\vec{y}}^{T}\mat{C}\vec{y} & \text{if } \vec{z}_3 = 1
    \end{cases}
\end{equation}
where $\mat{I}$ is the identity matrix, $\mat{U}$ is the unit matrix (all the elements are ones), $\mat{C}$ is a sparse stochastic matrix with $\mathrm{tr}(\mat{C})=0$ and $\mat{C}_{ij}$ denotes the confusion probability between classes $i$ and $j$. Then we can derive from Figure~\ref{fig:noise_pgm} the joint distribution of $\tilde{\vec{y}}, \vec{y}$ and $\vec{z}$ conditioning on $\vec{x}$:
\begin{equation} \label{eq:joint_distribution}
  p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}) = p(\tilde{\vec{y}} | \vec{y}, \vec{z}) p(\vec{y} | \vec{x}) p(\vec{z} | \vec{x})
\end{equation}

While the class label probability distribution $p(\vec{y} | \vec{x})$ is comprehensible, the semantic meaning of $p(\vec{z} | \vec{x})$ needs extra clarification: it represents how confusing the image content is. Specific to our clothing classification problem, $p(\vec{z} | \vec{x})$ can be affected by different factors, including background clutter, image resolution, the style and material of the clothes. Some examples are shown in Figure~\ref{fig:compare_noise_level}.

@@

To illustrate the relations between noisy and true labels, we derive their conditional probability from Eq~\ref{eq:joint_distribution}
\begin{equation}
    p(\tilde{\vec{y}} | \vec{y}, \vec{x}) = \sum_{\vec{z}} p(\tilde{\vec{y}}, \vec{z} | \vec{y}, \vec{x}) = \sum_{\vec{z}} p(\tilde{\vec{y}} | \vec{y}, \vec{z}) p(\vec{z} | \vec{x})
\end{equation}
which can be interpreted as a mixture model. Given an input image $\vec{x}$, the conditional probability $p(\vec{z} | \vec{x})$ can be seen as the prior of each mixture component. This makes a key difference between our work and~\cite{sukhbaatar2014learning}, where they assume $\tilde{\vec{y}}$ is conditionally independent with $\vec{x}$ if $\vec{y}$ is given. All the images share a same noise model in~\cite{sukhbaatar2014learning}, while in our approach each data sample has its own.

\subsection{Learning the Parameters} % (fold)
\label{sub:learning_the_parameters}
We exploit two CNNs to model $p(\vec{y} | \vec{x})$ and $p(\vec{z} | \vec{x})$ separately. Denote the parameter set of each CNN by $\theta_1$ and $\theta_2$. Our goal is to find the optimal $\theta = \theta_1 \cup \theta_2$ that maximize the incomplete log-likelihood $\log p(\tilde{\vec{y}} | \vec{x}; \theta)$. The EM algorithm is used to iteratively solve this problem.

For any probability distribution $q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x})$, we can derive a lower bound of the incomplete log-likelihood,
\begin{equation}
\begin{aligned}
    \log p(\tilde{\vec{y}} | \vec{x}; \theta) &= \log \sum_{\vec{y},\vec{z}} p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta) \\
    &\geq \sum_{\vec{y},\vec{z}} q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) \log \frac{p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta)}{q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x})}
\end{aligned}
\end{equation}
\textbf{E-Step} The difference between $\log p(\tilde{\vec{y}} | \vec{x}; \theta)$ and its lower bound is the Kullback-Leibler divergence $\mathrm{KL}\left(q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) || p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta)\right)$, which is equal to zero if and only if $q(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}) = p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta)$. Therefore, in each iteration $t$, we first compute the posterior of latent variables using the current parameters $\theta^{(t)}$:
\begin{multline} \label{eq:posterior}
    p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) = \frac{p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta^{(t)})}{p(\tilde{\vec{y}} | \vec{x}; \theta^{(t)})} \\
    = \frac{p(\tilde{\vec{y}}|\vec{y},\vec{z};\theta^{(t)}) p(\vec{y}|\vec{x};\theta^{(t)}) p(\vec{z}|\vec{x};\theta^{(t)})}{\sum_{\vec{y}',\vec{z}'} p(\tilde{\vec{y}}|\vec{y}',\vec{z}';\theta^{(t)}) p(\vec{y}'|\vec{x};\theta^{(t)}) p(\vec{z}'|\vec{x};\theta^{(t)})}
\end{multline}
Then the expected complete log-likelihood can be written as
\begin{equation}
    Q(\theta; \theta^{(t)}) = \sum_{\vec{y},\vec{z}} p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \log p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta)
\end{equation}
\textbf{M-Step} We exploit two CNNs to model the probability $p(\vec{y}|\vec{x};\theta_1)$ and $p(\vec{z}|\vec{x};\theta_2)$, respectively. Thus the gradient of $Q$ \wrt $\theta$ can be decoupled into two parts:
\begin{multline} \label{eq:gradient}
    \frac{\partial Q}{\partial \theta} = \sum_{\vec{y},\vec{z}} p(\vec{y}, \vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta} \log p(\tilde{\vec{y}}, \vec{y}, \vec{z} | \vec{x}; \theta) \\
    = \sum_{\vec{y}} p(\vec{y} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta_1} \log p(\vec{y}|\vec{x};\theta_1) + \\
      \sum_{\vec{z}} p(\vec{z} | \tilde{\vec{y}}, \vec{x}; \theta^{(t)}) \frac{\partial}{\partial \theta_2} \log p(\vec{z}|\vec{x};\theta_2)
\end{multline}

The M-Step above is equivalent to minimizing the cross entropy between the estimated ground truth distribution and the prediction of the classifier.
% subsection learning_the_parameters (end)

\subsection{Estimating Matrix $\mat{C}$} % (fold)
\label{sub:confusion_estimation}
Notice that we do not set parameters to the conditional probability $p(\tilde{\vec{y}}|\vec{y},\vec{z})$ in Eq~\eqref{eq:likelihood} and keep it unchanged during the learning process. Because without other regularizations, learning all the three parts in Eq~\eqref{eq:joint_distribution} could lead to trivial solutions. For example, the network will always predict $\vec{y}_1 = 1$, $\vec{z}_3 = 1$, and the matrix $\mat{C}$ is learned to make $\mat{C}_{1i} = 1$ for all $i$. To avoid such degeneration, we estimate $\mat{C}$ on a relatively small dataset $\mathcal{D}_c = \{(\vec{x},\vec{y},\tilde{\vec{y}})\}_N$, where we have $N$ images with both clean and noisy labels. As prior information about $\vec{z}$ is not available, we solve the following optimization problem:
\begin{equation} \label{eq:opt_all}
    \max_{\mat{C},\vec{z}^{(1)},\cdots,\vec{z}^{(N)}} \sum_{i=1}^N \log p(\tilde{\vec{y}}^{(i)}|\vec{y}^{(i)},\vec{z}^{(i)})
\end{equation}
Obviously, sample $i$ contributes nothing to the optimal $\mat{C}^*$ if $\vec{y}^{(i)}$ and $\tilde{\vec{y}}^{(i)}$ are equal. So that we discard those samples and reinterpret the problem in another form by exploiting Eq~\ref{eq:likelihood}:
\begin{equation}
\begin{aligned}
\max_{\mat{C},\vec{t}} \quad & E=\sum_{i=1}^{N'} \log \alpha^{\vec{t}_i} + \log (\tilde{\vec{y}}^{(i)T} \mat{C} \vec{y}^{(i)})^{1-\vec{t}_i}\\
\text{subject to} \quad & \mat{C} \text{ is a stochastic matrix of size } L\times L \\
                    & \vec{t} \in \{0,1\}^{N'}
\end{aligned}
\end{equation}
where $\alpha=\frac{1}{L-1}$ and $N'$ is the number of remaining samples. The semantic meaning of the above formulation is that we need to assign each $(\vec{y},\tilde{\vec{y}})$ pair the optimal noise type, while finding the optimal $\mat{C}$ simultaneously.
